Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-04 WSSSPE1_report/WSSSPE1_summary_report.tex
Added one additional attendee I was aware of
footnote'd nsf vision link

The First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)\footnote{\url{http://wssspe.researchcomputing.org.uk/WSSSPE1}} was held on Sunday, 17 November 2013, in conjunction with the 2013 International Conference for High Performance Computing, Networking, Storage and Analysis (SC13)\footnote{\url{http://sc13.supercomputing.org}}.

Because progress in scientific research is dependent on the quality and accessibility
of software at all levels, it is now critical to address many challenges related to the
development, deployment, and maintenance of reusable software. In addition, it is
essential that scientists, researchers, and students are able to learn and adopt
software-related skills and methodologies. Established researchers are already
acquiring some of these skills, and in particular a specialized class of software
developers is emerging in academic environments who are an integral and embedded
part of successful research teams. The WSSSPE workshop was intended to provide a
forum for discussion of the challenges, including both positions and experiences.
The short papers and discussion have been archived to provide a basis for continued
discussion, and the workshop has fed into the collaborative writing of this journal publication.
Additional papers based on extended versions of workshop submissions are expected.
The level of interest in the workshop has led the organizers, working with some of the submitters
and some of the attendees, to plan two additional workshops (\todo{details}) and to turn
the workshop website into a community website that can be used as a focus for further
discussion and progress.

@@

\todo{mention arXiv:1311.3523 and explain what's different here}

\section{Workshop Process and Agenda}

@@

\begin{quote}
In practice, scientific software activities are part of an ecosystem where key roles are held by 
developers, users, and funders. All three groups supply resources to the ecosystem, as well as 
requirements that bound it. Roughly following the example of NSF's Vision and Strategy for Software 
(\url{http://www.nsf.gov/publications/pub_summ.jsp?ods_key=nsf12113})~\cite{NSF_software_vision}, 
the ecosystem may be viewed as having challenges related to:

@@

\section{Keynotes}

@@

He noted that in places, development in science has improved thanks to
open source and Github, but for the most part remains arcane. He argues
that there is a lot we can learn from the app-store model. He
mentioned BioJava\footnote{\url{http://biojava.org/wiki/Main_Page}} and
OpenScienceDataCloud \footnote{\url{https://www.opensciencedatacloud.org/}} as
distribution channels.

@@

\section{Developing and Supporting Software}

@@

\section{Policy}

@@

\section{Communities}

@@

\section{Other Discussion}

@@

\section{Cross-cutting Issues}

@@

\section{Illustrative Use Cases}

@@

\section{Conclusions}

@@

>>>>

The First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1, \url{http://wssspe.researchcomputing.org.uk/WSSSPE1}) was held on Sunday, 17 November 2013, in conjunction with the 2013 International Conference for High Performance Computing, Networking, Storage and Analysis (SC13, \url{http://sc13.supercomputing.org}).

Because progress in scientific research is dependent on the quality and accessibility
of software at all levels, it is now critical to address many challenges related to the
development, deployment, and maintenance of reusable software. In addition, it is
essential that scientists, researchers, and students are able to learn and adopt
software-related skills and methodologies. Established researchers are already
acquiring some of these skills, and in particular a specialized class of software
developers is emerging in academic environments who are an integral and embedded
part of successful research teams. The WSSSPE workshop was intended to provide a
forum for discussion of the challenges, including both positions and experiences.
The short papers and discussion have been archived to provide a basis for continued
discussion, and the workshop has fed into the collaborative writing of this journal publication.
Additional papers based on extended versions of workshop submissions are expected.
The level of interest in the workshop has led the organizers, working with some of the submitters
and some of the attendees, to plan two additional workshops (at the 2014 SciPy and SC conferences) and to turn
the workshop website into a community website that can be used as a focus for further
discussion and progress.

Before the WSSSPE1 workshop took place, the organizers self-published a paper~\cite{WSSSPE1-pre-report} to document the process of organizing and advertising the workshop, collecting and reviewing the papers, and putting together the agenda.  Section~\ref{sec:process} is based on that paper.  The remainder of this paper is based on the workshop itself, as documented in a set of collaborative notes taken during the workshop~\cite{WSSSPE1-google-notes}, including discussion about the keynote presentations (\S\ref{sec:keynotes}), the three panels (\S\ref{sec:devel}-\ref{sec:community}), other discussion (\S\ref{sec:other}), cross-cutting issues (\S\ref{sec:cross-cutting}), use cases (\S\ref{sec:use-cases}), and conclusions (\S\ref{sec:conclusions}).

@@

\section{Workshop Process and Agenda} \label{sec:process}

@@

\begin{quote}
In practice, scientific software activities are part of an ecosystem where key roles are held by 
developers, users, and funders. All three groups supply resources to the ecosystem, as well as 
requirements that bound it. Roughly following the example of NSF's Vision and Strategy for Software 
\footnote{\url{http://www.nsf.gov/publications/pub_summ.jsp?ods_key=nsf12113}}~\cite{NSF_software_vision}, 
the ecosystem may be viewed as having challenges related to:

@@

\section{Keynotes} \label{sec:keynotes}

@@

He noted that in places, development in science has improved thanks to
open source and Github, but for the most part remains arcane. He argues
that there is a lot we can learn from the app-store model. He
mentioned BioJava\footnote{\url{http://biojava.org/wiki/Main_Page}} and
OpenScienceDataCloud\footnote{\url{https://www.opensciencedatacloud.org/}} as
distribution channels.

@@

\section{Developing and Supporting Software} \label{sec:devel}

@@

\section{Policy} \label{sec:policy}

@@

\section{Communities} \label{sec:community}

@@

\section{Other Discussion} \label{sec:other}

@@

\section{Cross-cutting Issues}  \label{sec:cross-cutting}

@@

\section{Illustrative Use Cases} \label{sec:use-cases}

@@

\section{Conclusions} \label{sec:conclusions}

@@

Patrick O'Leary
